### 2017-02-08 Icebergs_only.tex
Adding iceberg only version of the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified July 2011
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \defcommand, or \renewcommand.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format: two column (default)
% and draft.
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER

% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft,jgrga]{AGUTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OPTIONAL:
% To produce a two-columned version:
% \documentclass[grl]{AGUTeX}

% Two-columned format can be used to estimate the number of pages
% for the final published PDF.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OPTIONAL:
% To create numbered lines:

% If you don't already have lineno.sty, you can download it from
% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
% (or search the internet for lineno.sty ctan), available at TeX
%Archive Network (CTAN).
% Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno}
% and \linenumbers*[1]command, below:
% To activate the commands, uncomment \usepackage{lineno}
% and \linenumbers*[1]command, below:
\usepackage{amsmath, amsthm, amssymb}
 \usepackage{lineno}
 \linenumbers*[1]

 %\usepackage{lineno}
 %\linenumbers*[1]

%  To add line numbers to lines with equations:

%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
% When submitting articles through the GEMS system:
% COMMENT OUT ANY COMMANDS THAT INCLUDE GRAPHICS.
% (See FIGURES section near the end of the file.)
%
% DO NOT USE \psfrag or \subfigure commands.
%
%  Figures and tables should be placed AT THE END OF THE ARTICLE,
%  after the references.
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
 \usepackage{graphicx}
 \usepackage{ulem}
\usepackage{color}
  %
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
  \setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
\authorrunninghead{STERN ET AL.}

% Shorter version of title entered in capital letters:
%\titlerunninghead{Remote and local affects of large Antarctic icebergs}
%\titlerunninghead{Effects of iceberg calving size}
\titlerunninghead{Modeling tabular icebergs}

% Author mailing address: please repeat this command for
% each author and alphabetize authors:

\authoraddr{A. A. Stern, Geophysical Fluid Dynamics Laboratory, Princeton University}

\authoraddr{A. Adcroft, Geophysical Fluid Dynamics Laboratory, Princeton University}

\authoraddr{O. Sergienko, Geophysical Fluid Dynamics Laboratory, Princeton University}

\authoraddr{G. Marquez, Geophysical Fluid Dynamics Laboratory, Princeton University}

\authoraddr{R. Hallberg, Geophysical Fluid Dynamics Laboratory, Princeton University}

%
\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

%\title{Remote and local effects of large Antarctic icebergs}
%OR\\
%\title{The effect of calving-size distribution on sea-ice formation around Antarctica}
\title{Modeling tabular icebergs coupled to an ocean model}
%\title{The effects of iceberg calving-size distribution in a global climate model}
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

%Use \author{\altaffilmark{}} and \altaffiltext{}

 %\altaffilmark will produce footnote;
% matching \altaffiltext will appear at bottom of page.

 \authors{A.A. Stern,\altaffilmark{1} , A. Adcroft\altaffilmark{1}
   and O. Sergienko \altaffilmark{1} , G. Marquez\altaffilmark{1}, R. Hallberg\altaffilmark{1}}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%

% >> Do NOT include any \begin...\end commands within
% >> the body of the abstract.

\begin{abstract}
The calving of giant icebergs from the Antarctic ice shelves accounts for approximately half of Antarctic ice-shelf decay. After calving these tabular icebergs drift across large distances, altering regional ocean circulation, bottom water formation and local rates biological primary productivity. Freshwater fluxes from sub-marine melting can affect large-scale sea-ice distributions, which feedback onto the greater climate system.
However, despite their importance, the current generation of numerical models is unable to represent giant tabular icebergs a physically realistic way.
In this study we develop novel method to model giant tabular icebergs. Tabular icebergs are constructed out of Lagrangian elements which are bonded together by numerical bonds. 
These numerical bonds hold the ice element together and allow the individual elements to move as a unit. By breaking these bonds, we can simulate the  ocean response to giant iceberg detaching away from an ice shelf (also constructed from Lagrangian elements) or the fracturing of a large iceberg into smaller pieces.  This paper provides a description and demonstration of the Lagrangian tabular-iceberg model and discusses a number of the technical elements involved its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg breaking away an idealized ice shelf.
\end{abstract}
%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%
% The body of the article must start with a \begin{article} command
%
\clearpage
\vspace{10mm}
\textbf{Key Points:}\\

% \end{article} must follow the references section, before the figures
%  and tables.
\begin{article}
%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%
\section{Introduction}
%\subsection{Key points for writing introduction}
%
%1) Physical processes that affect the ocean circulation around Antarctica (involving ice shelf cavities /tabular icebergs). 
%
%2) Why is it important that we are about to model these regions correctly, and tabular icebergs?
%
%3) What are the significant efforts that have been made thus far towards correctly modeling ice shelf cavities / Antarctic fresh water fluxes.
%
%4) What is missing - and what we intend to add
%
%5) Similar things which have been done in the past (if there are any)
%
%6) Short summary and paper description.
%
%\subsection{Introduction}

The Antarctic ice shelves are characterized by large infrequent calving events where massive pieces of the ice shelves break off, to create giant tabular icebergs. Observational estimates suggest that over the past 30 years approximately half of Antarctic ice-shelf decay is due to iceberg calving, while the other half occurs through ice-shelf melting \citep{Depoorter2013, Rignot2013}. After calving, icebergs slowly drift away from their mother glaciers, often becoming stuck in sea ice, or become grounded along the Antarctic coast \citep{Lichey2001, Dowdeswell2007}. Giant tabular icebergs extend deep into the water column, and have to potential to disrupt ocean circulation patterns for months or even years after calving \citep{Robinson2012, Stern2015}. Iceberg sizes appear to fit a power law distribution, with the majority of icebergs being small (L$<$ 1 km), while the far less numerous giant tabular icebergs (L $>$ 5km) account for more than 90$\%$ of the icebergs mass \citep{Tournadre2016}.

Freshwater fluxes from iceberg melt impact ocean hydrography, influencing sea-ice production and bottom-water formation \citep{Arrigo2002, Robinson2012, Nicholls2009}. The long distances traveled by giant icebergs before melting means that their meltwater impact is often felt hundreds of kilometers away from their calving origins \citep{Stern2016}. Melt water injection (and the accompanying upwelling) from icebergs can also influence biological productivity by providing nutrients to the surface ocean or changing sea ice conditions \citep{Arrigo2002, Vernet2012, Biddle2015}. The increased productivity associated with free-floating tabular icebergs has been linked with local increases in ocean carbon uptake, potentially large enough to be a significant fraction of the Southern Ocean carbon sequestration \citep{Smith2007}.
%either increasing it by providing nutrients to the surface ocean, or decreasing by changing sea ice conditions

In recent years, there has been an increased interest in iceberg drift and decay. This surge of interest has been driven by (i) the need to understand polar freshwater cycles in order to create realistic climate forecasts and sea level projections \citep{Silva2006, Shepherd2007, Rignot2013}, (ii) the increased use of high-latitudes iceberg-filled waters for shipping lanes and offshore hydrocarbon exploration \citep{Pizzolato2014, Unger2014, Henderson2016}. The increased interest in icebergs has led to the development of numerical models of iceberg drift and decay \citep{Mountain1980, Bigg1997, Gladstone2001, Kubat2005}, some of which have been included in global General Circulation Models \citep{Martin2010, Marsh2015}. These iceberg models treat icebergs as Lagrangian point particles, which are advected by the flow, and melt according to certain parameterizations for icebergs melt. Since icebergs are treated as point particles, the iceberg models are mostly suitable for modeling smaller icebergs drifting across large distances, and as such iceberg models have mostly been used to represent icebergs smaller than 3.5km on a global scale \citep{Jongma2009, Martin2010, Marsh2015}.
These models are not suitable for modeling larger tabular icebergs (or smaller icebergs at fine scales), where the size and structure of the iceberg may be an important feature in determining their drift and decay \citep{Stern2016}. They also are not suitable for studying the local effects that icebergs have on the surrounding ocean. For this reason, tabular icebergs (L $>$ 5km) are currently not represented in the iceberg models used in climate models, even though observations suggest that tabular icebergs account for  the vast majority of the total Southern Hemisphere icebergs mass \citep{Tournadre2016}. Similarly, operational iceberg models are often inadequate for projecting iceberg drift at fine scales \citep{Turnbull2015}.

The goal of this study is to develop a new iceberg model where tabular icebergs are explicitly resolved in the ocean. In this model, icebergs are no longer treated as point particles that feel the ocean at one point, but rather the icebergs have structure, depress the ocean surface, and can feel the ocean across multiple ocean grid cells. To do this, we create tabular icebergs out of many Lagrangian elements which are `bonded' together by numerical bonds. The numerical bonds hold the ice elements together and allow a collection elements to move as a unit. By manually breaking these bonds, we can simulate iceberg calving, allowing us to study the ocean response to giant iceberg detaching from an ice shelf or the fracturing of a large iceberg into smaller pieces.

The discrete element model (DEM) developed in this study is referred to at the Lagrangian Tabular Iceberg Model (LTAM). Section 2 gives a description of many of the key aspects of the LTAM. Since this model is a new approach to modeling iceberg, we present many of the technical elements involved in constructing the model. In Section 3 and 4, we demonstrate the capabilities of the model by simulating a tabular iceberg detaching from an idealized ice shelf.

%A prerequisite of a complete model of an ice shelf cavity with a calving ice shelf is an understanding of the physical processes involved in iceberg calving. The importance of this problem for projecting sea level has meant that the search for a ``calving law" has received much attention in recent years. However, even with a theory of ice shelf breaking, there are still significant technical barriers preventing us from modeling icebergs calving away from ice shelves in a dynamically consistent way. In this paper, we d

%This work bears some similarity to [Basis et al], who used a DEM model to model the generation and propagation of cracks within an ice shelf. The present study is distinct from [Basis et al] in that (i) we  focus on the ocean's response to iceberg calving and thus our DEM model is coupled to a fully dynamic ocean model, (ii) the element in our DEM model are of scale > 1 km, while those in Basis were much smaller (1 meter), since we focus on the large scale oceanography, rather than the small scale fracture mechanics. Finally, we do not model the breaking in iceberg bonds and ice dynamics. However, including bonds between ice elements in our model means that modeling ice shelf breaking is a natural extension to this work.

%This same constraint is present in previous iceberg models, and has meant that those models have only been able to represent smaller icebergs (e.g. icebergs with area less than $3.5km^{2}$ in \citet{Gladstone2001, Martin2010}). However, as mentioned above, satellite observations of iceberg size show that more than 80(?) percent of Southern Hemispheric icebergs mass is stored in icebergs with areas greater than $3.5km^{2}$  \citep{Tournadre2016}, which motivates the including of larger icebergs in climate models.

\section{Model description}
%, which can be used to model small icebergs, tabular icebergs, sea ice and ice shelf cavities
%In this section we describe many key elements of the LTAM.  
The LTAM is a Lagrangian particle-based model (or DEM model) in that the objects of the model are Lagrangian elements. Each element represents a mass of ice which is floating in the ocean, and has a position, velocity, mass, and a set of dimensions, which can evolve in time. The motion of each element is determined by a momentum equation which is solved in the (Lagrangian) reference frame of the element. The elements are forced by oceanic and atmospheric forces, which are provided by the user, or are determined by coupling the LTAM to an ocean/atmosphere model. The elements also interact with one another and can be bonded together to form larger structures. The angular momentum of the elements is not model explicitly, instead rotational motion of larger structures emerge as a consequence of bond orientation. In different contexts, the LTAM elements can be thought to represent individual icebergs, sea ice flows, or, when the elements are bonded together, they can represent larger structures such as tabular icebergs or ice shelves.
%In this latter case, the LTAM model can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

The LTAM model is developed on the code base of an existing iceberg drift model \citep{Martin2010, Stern2016}. When run with the correct set of runtime flags, the model runs as a traditional iceberg drift model.
%
%In this section, we discuss the equations of motion governing the element movement and the thermodynamic equations governing the decay of the elements. The also discuss the interactive forces

\subsection{Equations of motion}
The elements drift in the ocean, forced by atmosphere, ocean and sea-ice drag forces, as well as the Coriolis force and a force due to the sea surface slope.  When these ice elements move alone (without interactions with other elements), they can be thought of as representing individual (or clusters of) small icebergs, and follow the same equations described in the iceberg drift model of \citet{Martin2010} (based on the equations outlined in \citet{Bigg1997, Gladstone2001}).

In addition to the environmental forces, the elements in the LTAM model experience interactive forces due to the presence of other elements. Two types of interactive forces are included between elements. The first force is repulsive force which is applied to elements to prevent them from moving too close to one another. This repulsive force prevents icebergs from piling up on top of one another. The second interactive force is a force due to numerical `bonds', and is only applied if two elements are `bonded' together. When two elements are bonded, each element feels an attractive force that prevents the elements from moving too far apart from one another. The interactive forces between two bonded elements are defined such that in the absence of other forces the elements come to rest adjacent to one another, with no overlap of the iceberg areas.

The momentum equation for each element is given by 
\begin{equation}
\label{Iceberg_mom_eq}
M \frac{D \vec{u}}{Dt}=\vec{F}_{A} + \vec{F}_{W}+ \vec{F}_{R}+\vec{F}_{C} +\vec{F}_{SS}+\vec{F}_{SI}+\vec{F}_{IA} ,
\end{equation}
where $\frac{D}{Dt}$ is the total (Lagrangian) derivative, $M$ is the mass of the element, $\vec{u}$ is the velocity of the element, and the terms on the right hand side give the forces on the element due to air drag $(\vec{F}_{A})$, water drag $(\vec{F}_{W})$, sea ice drag $(\vec{F}_{SI})$, Coriolis force $(\vec{F}_{C})$, wave radiation force $(\vec{F}_{R})$, sea surface slope $(\vec{F}_{SS})$, and interactions with other elements ($\vec{F}_{IA}$). The environmental forces are the same as those presented in \citet{Martin2010}, and are provided for completeness in Appendix A.  The details of the interactive forces are provided in below.

\subsection{Interactive Forces}
The interactive force on an element is calculated by adding together the interactions with all other elements, such that the interactive force on element $i$, $(\vec{F}_{IA})_{i}$ is given by:
\begin{equation}
\label{F_interaction_i}
(F_{IA})_{i}= \Sigma_{j \neq i} (F_{IA})_{ij},
\end{equation}
where $(F_{IA})_{ij}$ is the force on element $i$ by element $j$. Both bonded and repulsive interactions are modeled using elastic stresses with frictional damping. The elastic component of the force is a function of the distance between the two elements, while the  frictional damping force depends on the relative velocity of the two elements.

To describe the forces between two elements, we begin by introducing some notation.  Let $\vec{x_{i}}$, $\vec{x_{j}}$ be the positions of elements i and j. The distance between elements i and j is
\begin{equation}
\label{distance}
d_{ij}= | \vec{x_{i}} - \vec{x_{j}}|.
\end{equation}
When calculating the interactive forces between elements, the elements are assumed to be circular. We define the interaction diameter of an element by 
\begin{equation}
\label{Diameter}
D_{i}=2 \sqrt{\frac{A_{i}}{\pi}},
\end{equation}
where $A_{i}$ is the planar surface area of element i. Using this, we define the critical interactive length scale, 
\begin{equation}
L_{ij}=\frac{D_{i} + D_{j}}{2},
\label{eq:Interaction_crit}
\end{equation}
which governs interactions between elements i and j. 
Repulsive forces are only applied when $d_{i,j} \textless L_{i,j}$, while for $d_{i,j} \textgreater L_{i,j}$ attractive bonded forces are applied when a bond exists between element i and j.  Bond and repulsive forces are designed such that in the absence of other forces, bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{i,j}$.
% (see schematic in Figure \ref{fig:Interactive_forces}).

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements i and j are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F_{IA}})_{ij}$ on an element i by an element j is given by
\begin{equation}
\label{F_interactione}
(\vec{F_{IA}})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F_{e}})_{ij} + (\vec{F_{d}})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F_{e}})_{ij} + (\vec{F_{d}})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements i and j.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) T_{i,j} \vec{r},
\end{equation}
where $\vec{r}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element i and j, and $\kappa_{e}$ is the spring constant, and $T_{i,j}$ is the minimum of the thickness of elements i, j. We choose $(\vec{F_{b}})_{ij}$ proportional to $T_{i,j}$ (the minimum thickness of elements i and j) so the formulation of interactive forces has the desired property that it obeys Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum thickness is preferred to the average thickness, since this means that the for two bonded elements a fixed distance apart, acceleration due to elastic forces is bounded, even when the thickness of one of the elements approaches zero. 
%
%Since ice elements are able to melt entirely, the repulsive forces between elements need to be carefully defined so that they are stable even for very small elements with mass approaching zero. We define the elastic repulsive force $(F_{r})_{ij}$ between elements i and j as
%\begin{equation}
%\label{F_elastic}
%(F_{r})_{ij}= -\kappa_{r} A_{o}(i,j) T_{i,j}\vec{r},
%\end{equation}
%where $A_{o}(i,j)$ is the overlap area of the two circles centered at $\vec{x_{i}}$ and $\vec{x_{j}}$, with diameter $D_{i}$ and $D_{j}$ (see Figure \ref{fig:Interactive_forces}), respectively, and $\kappa_{r}$ is the repulsive spring constant. We choose $(F_{r})_{ij}$ to be proportional to the overlap area between the interactive circles surrounding each element so that the acceleration due to this repulsive force is bounded, even as the element mass approaches zero.   This consideration is not needed for the bonded force, since element bonds will be broken when elements are smaller than a threshold size. 
%[Should we use a bonded force given by  $(F_{b})_{ij}= -\kappa_{b}  \bigg( d_{i,j}-L_{i,j}) \bigg)^{2} T_{i,j} \vec{r}$ rather? Come back to this.]

The frictional damping force has components that damp both the relative radial velocity and relative transverse velocities of the two elements. If $\vec{r}^{\perp}$ is the direction vector perpendicular to $\vec{r}$, and $P_{\vec{r}}$ and $P_{\vec{r^{\perp}}}$ are the projection matrices that project onto $\vec{r}$ and $\vec{r}^{\perp}$ respectively, then the frictional damping force is given by
\begin{equation}
\label{F_damping} 
%(F_{d})_{ij}=   -C_{r}|(u_{i} -u_{j}) \cdot \vec{r}   | \vec{r}  - C_{t}|(u_{i} -u_{j})\cdot  \vec{r}^{\perp} | \vec{r}^{\perp})
(\vec{F_{d}})_{ij}=   \Bigg (-c_{r} P_{\vec{r}}  - c_{t} P_{\vec{r}^{\perp}} \Bigg ) \cdot (\vec{u_{i}} -\vec{u_{j}} ) 
\end{equation}
Here $c_{r}$ and $c_{t}$ are the radial and transverse drag coefficients. For the simulation below, we set $c_{r}=2\sqrt{\kappa_{e} }$ and  $c_{t}=\frac{1}{4}c_{r}$ so that the radial elastic force is critically damped, and the transverse damping is sub critical. The damping forces are implemented using an implicit time stepping scheme, to avoid stability issues for very small elements (details found in Appendix B).

The effectiveness of the repulsive forces can be seen in Figure \ref{fig:Regular_Uncoupled_sim}, which shows an uncoupled (ice only) simulation where ice elements drift westward into a bay, and eventually come to rest with minimal overlap between elements. The effectiveness of the numerical bonds is demonstrated in Figure \ref{fig:Uncoupled_sim}, where tabular icebergs (constructed from many ice elements bonded together) and individual icebergs (unbonded elements) drift together towards a convex coast line. When the tabular icebergs arrive at the coast, they bump into the coastline and begin to rotate, influencing the paths of the other icebergs. In this example we see that modeling large structures using small elements bonded together, allows us to achieve large-scale structure and rotational motion, without having to include an equation for the angular momentum of the elements (as discussed in \citet{Jakobsen2001}). Animations of these simulations can be found in the supplementary materials.

\subsection{Element geometry and packing}
For purposes of packing, we assume that elements have surface areas which are shaped as equally-sized regular hexagons (note that the elements are assumed to be circular for proposes of interactions, but are assumed to be hexagonal for packing purposes). When packing these elements together, the hexagonal elements are initially arranged in a staggered lattice, with each element bonded to the adjacent elements (see Figure  \ref{fig:Hex_intersections}a). In this arrangement, each element (away from the edges) is bonded to six other elements. The bonds between elements form a pattern of equilateral triangles, which give the larger structure rigidity. The circular shape of elements (used for interactions) is inscribed within the hexagonal shape used for packing (Figure  \ref{fig:Hex_intersections}a). The centers of adjacent elements are initially separated by a distance $d_{i,j}=L_{i,j}=2A_{p}$, where $A_{p}$ is the length the apothems of the hexagons.

Some experiments were also performed using rectangular elements, arranged in a regular (non-staggered) lattice. In this case, each element forms four bonds with adjacent elements. However, the resultant structures were found to be much less rigid and tended to collapse when sufficient forces was applied. For this reason, hexagonal elements are used here.

\subsection{Ocean-ice and ice-ocean coupling}
The LTAM model is coupled to the ocean model via a two-way synchronous coupling, meaning that ocean model fields are passed to the LTAM model and the LTAM model fields are passed back to the ocean model at every time step. Passing fields between the two models involves interpolating fields between the Eulerian grid of the ocean model and the `Lagrangian grid' of the LTAM model (i.e.: onto the ice elements).

The coupling from the ocean model to the LTAM model is straight forward: at every time step, the ocean mixed layer temperature, salinity, velocity and sea-ice concentration are passed from the ocean model to the LTAM model, to be used in the momentum and thermodynamic equations of the ice elements. Within the LTAM model, the ocean model fields are interpolated onto the Lagrangian grid using a bilinear interpolation scheme. The LTAM model is not sensitive to the chosen interpolation scheme.

%More care needs to be taken when passing fields back from the LTAM model to the ocean model. 
The coupling in the other direction, from the LTAM model to the ocean model, is more complex.
The LTAM model influences the ocean by: (i) applying a pressure to the ocean surface, (ii) applying a heat, salinity and mass flux to the ocean, associated with ice melting, and (iii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice. Fields from the LTAM model are interpolated from the Lagrangian grid to the Eulerian ocean grid before they are passed to the ocean model. Since the LTAM model applies large pressures to the ocean surface, the ocean model is sensitive to the interpolation scheme used to map the LTAM fields to the ocean grid. Sudden jumps to the ocean surface pressure can trigger tsunamis within the ocean model, making the ocean model become unstable.

The interpolation of the LTAM fields onto the ocean grid is done in a way that is consistent with the shape of the elements in the LTAM model (see Section 2.3). Fields are `spread' to the ocean model grid by exactly calculating what fraction of an element's surface area lies in a particular grid box, and dividing the field in proportion to this fraction.
%While the position of a particular element might be inside one ocean grid cell, the implied shape of the element will often intersect other ocean grid cells. 
For example, consider a hexagonal element in the LTAM model, which is positioned such that it intersects four ocean grid cells (Figure \ref{fig:Hex_intersections}b). In this situation, the mass of the element is divided between these four ocean cells in proportion to the overlap area between the hexagonal element and the grid cell (this fraction is shown by the colors in \ref{fig:Hex_intersections}b). An advantage of this approach is that there are no jumps in pressure as an element moves from one grid cell to another.

The numerical calculation of the intersection between hexagons and the ocean grid is simplified by dividing the hexagon into 6 equilateral triangles. This method allows for the intersection to be found even when the hexagon is not aligned with the grid. %The LTAM model can be run using hexagonal or rectangular elements.

The field-spreading scheme is coded with the restriction that an element's area can only intersect a maximum of four ocean grid cells at a time. A consequence of this is that this sets a limit on the maximum size of elements that can be represented using this model, i.e.: the longest horizontal dimension of an ice element must be smaller than the ocean grid spacing. Larger ice structures are constructed by bonding together smaller elements.

\subsection{Thermodynamics}
The ice elements decay according to a number of melt parameterizations.  As the ice elements melt, their mass decreases, and the appropriate salt, mass and heat fluxes are passed to the ocean. In this section we described the melt parametrization for bonded, unbonded and partially bonded elements.

As mentioned above, ice elements which do not interact with other elements are modeled identically to the point particle icebergs described in \citet{Martin2010}. These elements melt according to three semi-empirical parametrization for melt commonly used in previous iceberg studies \citep{Gladstone2001,Martin2010}.  Three types of iceberg melting are used: basal melt, $M_{b}$, melt due to wave erosion, $M_{e}$ and melt due to buoyant convection, $M_{v}$. $M_{e}$ and $M_{v}$ are applied to the sides of the ice element, while $M_{b}$ is applied at the ice element base. The details of $M_{b}$. $M_{v}$ and $M_{e}$ are given in Appendix A.

When multiple elements are bonded together to form larger structures, it is no longer appropriate to use the parameterizations for melt developed for individual point-particle icebergs. An element which is completely surrounded by other elements, is meant to represent a piece of ice in the middle of a large structure, and hence will not experience a melt at its sides due to wave erosion or buoyant convection. Also, the iceberg basal melt rate, $M_{b}$ described above is based on boundary layer theory of flow past a finite plate, and is only appropriate for basal surfaces where the distance from the leading edge is sufficiently small \citep{Eckert1950, Weeks1973}. For an element in the interior of large structures, the distance from the edge of the structure is large, and so using $M_{b}$ for the basal melt is not appropriate. Instead, the basal melt, $M_{s}$ is determined using the three equation model for basal melt, which is a typical melting parametrization beneath used beneath ice shelves \citep{Holland1999}.% (see Appendix A).

When using both individual elements and bonded elements in the same simulation, we determine which melt rate parameterizations to use based on the amount of bonds that each element has. An element which is in the center of a large structure will form the maximum number of bonds, while unbonded elements form zero bonds. If maximum number of bonds that an element can form (given the shape of the element) is $N_{max}$, and the number bonds that an element has is $N_{b}$, then the side melt and bottom melt for that element are given by
\begin{equation}
M_{side}=\frac{(N_{max}-N_{b})}{N_{max}}  (M_{v} +M_{e})
\end{equation}
and
\begin{equation}
M_{bottom}=\frac{(N_{max}-N_{b})}{N_{max}}M_{b} +\frac{N_{b}}{N_{max}}M_{s}
\end{equation}
respectively.  In this way, elements with no bonds, melt like point particle icebergs, elements at the center of large structures melt like ice shelves, and elements at the sides of large structures have a combination of iceberg side and basal melt, and ice-shelf melt.

\subsection{Algorithms and computational efficiency}
Including interactions between elements leads to an increase in the computational complexity of the model. In this subsection we comment on some of the algorithmic procedures that have been used to increase the computational efficiency.

\subsubsection{Interactions and Bonds}
At every time step, we calculate the force on each element due to interactions with every other element. In principle, this involves order $n^{2}$ operations (for n elements). However, since each element only has repulsive interactions with elements that are less than one ocean grid cell away, and each element only has bonded interactions with a small number of other elements, we are able to reduce the complexity of the system.

The complexity reduction is achieved by storing the element data in an efficient way that prevents having to search through all element pairs to check if they are close to one another or are bonded with one another. The data storage system works as follows: pointers to the memory structures containing each element are stored in linked list data structures, which allow elements to be added and removed from the lists easily without restructuring the entire list. Instead of using one list for all the elements on a processor (as was done in the original code \citep{Martin2010}), we use a separate linked list for each ocean grid cell. When an element moves between ocean grid cells, it is removed from its original list and added to the list corresponding to its new ocean grid cell. Since the area of elements has to be smaller than the area of an ocean grid cell, the critical interaction length scale (equation \ref{eq:Interaction_crit}) is less than the length of a grid cell. This means that elements only experience repulsive forces with elements in the same ocean grid cell, or in one of the 8 adjacent cells. Limiting the possible repulsive interactions to elements in these 9 linked lists substantially reduces the computational time needed to calculate the total interactive force.

Bonded interactions are handled differently. Each bond is assigned a piece of memory. Each ice element contains a linked list of each of its bonds (typically up to six bonds per element). At each time step, the code traverses the lists of bonded elements, and adds a bonded force corresponding to these bond. The bonded force is only applied if $d_{i,j} > L_{ij}$, to avoid double counting an interaction. Having a list of bonds stored with each element reduces the computational complexity of bonded interactions from order $n^{2}$ to order n. Handling bonded attractive forces separately to the repulsive and non-bonded forces means that we do not need to check whether two elements are bonded, which further increases the computational efficiency.

\subsubsection{Parallelization and halos}
The LTAM model runs on multiple processors in parallel. When elements move from an ocean cell on one processor to an ocean cell on a second processor, the memory has to be passed from one processor the next, added and removed to the appropriate lists and the memory has to be allocated and deallocated correctly. Element interactions across the edge of processors are handled using computational halos. A computational halo is a copy of the edge of a one processor which is appended to the edge of a second processor, so that the first processor can `see' the second processor during a time step. Before a time step, elements at the edges of each processor are copied onto the halos of adjacent processors so that they can be used in calculating the interactive forces. After each time step, these halos and removed, and the process is repeated. These halo updates are one of the most computationally expensive parts of the LTAM model.

Keeping track of pairs of bonded elements that move across a processor edge requires a lot of book keeping since bonds have to be severed and reconnected. Details of how the bonds are broken and reconnected across processor boundaries are provided in Appendix~C.

\subsubsection{Time stepping}
The ice elements in the LTAM model are advected using a semi-implicit velocity Verlet time-stepping scheme. The velocity verlet time stepping scheme is commonly used in DEM models in video games because it is computational efficient and has desirable stability properties \citep{Jakobsen2001}. This time stepping scheme was preferred to the Runge-Kutta  4, which was used in the iceberg model of \citet{Martin2010} since, the Verlet time stepping only requires one calculation of the interactive forces per time step (while the Runge-Kutta scheme requires the interactive forces to be calculated four times). Since the calculation of the interactive forces is one of the most computationally expensive part of the algorithm, the Verlet scheme leads to a significant increase in the speed of the model. We note that the Verlet scheme used in the LTAM contains a small modification of the original (fully explicit) velocity Verlet time stepping scheme in that damping terms are treated implicitly (which increases the numerical stability). The details of this time stepping schemed are outline in Appendix B.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section{Experiment Setup}
%The introduction of Lagrangian elements, numerical bond and appropriate interpolation schemes between the Eularian and Lagrangian grids (discussed in Section 2) means that we are now able to model ice shelf cavities which undergo large calving events, and create giant tabular icebergs.

%We demonstrate this capability in an experimental setup using an idealized ice shelf. For these experiments the LTAM model is coupled to the MOM6 ocean model.
% This section contains a description of the experimental setup and numerical experiments performed. The results of these simulations are presented in Section 4.

 %In the examples below, the LTAM model is firstly used in uncoupled mode to model the movement of freely drifting icebergs. Secondly, the coupled LTAM with static ice elements is used to model an idealized ice shelf cavity. Finally, the fully-coupled LTAM is used to model a tabular iceberg which calves off of this ice shelf.

%\subsection{Uncoupled simulation}
%We begin by demonstrating the LTAM model uncoupled to the ocean. For this experiment, we use a rectangular domain, $L_{x}=1000$ km long and $L_{y}=500$ km wide. Land points are arranged in a semi-circle, in the west of the domain, as shown by the black dots in Figure \ref{fig:Uncoupled_sim}.  Three hundred ice elements are initialized with diameters uniformly sampled from the interval [0 40km], and positioned randomly at positions with $x \leq 300$km. These ice elements can be thought to represent interacting sea ice flows or icebergs. Three `tabular icebergs' are initialized by arranging ice elements in a staggered lattices with adjacent elements bonded together, as described in Section 2.4. The three tabular icebergs contain 25, 16 and 4 elements, respectively. Each of the elements that make up the tabular icebergs has a diameter of 40km. All elements have an initial thickness of 10m.
%
%The simulation is forced by an imposed westward ocean current of u=0.1m/s. Forces due to sea surface slope, atmospheric drag, Coriolis and sea ice drag are set to zero. The initial setup is shown in Figure \ref{fig:Uncoupled_sim}a.

%\subsection{Coupled Ocean-Ice simulations}
%\subsection{Description of idealized ice shelf cavity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section{Experiment Setup}
The introduction of Lagrangian elements, numerical bonds and interpolation schemes between the Eularian and Lagrangian grids (discussed in Section 2) means that we now have the tools to model giant tabular icebergs submerged in the ocean. We demonstrate this capability by performing a simulation of a tabular iceberg drifting away from an idealized ice shelf. This simulation is performed using the LTAM model coupled to the MOM6 ocean model \citep{Hallberg2013}. 
%This section describes the experimental setup of these numerical experiments. The results are shown in Section 4.

 \subsection{Model configuration}
We initialized our simulation using the experimental setup created for the Marine Ice Ocean Modeling Inter-comparion Project (MISOMIP) \citep{Asay-Davis2016}.
The configuration consists of an idealized ice shelf in a rectangular domain. The domain is $L_{x}=80$km wide and $L_{y}=480$km long, and contains an ice shelf which is grounded on the south side of the domain and has an ice front at y=650 km. The ice thickness and bottom topography of this setup are shown in Figure \ref{fig:ISOMIP_mass_and_topog}a and \ref{fig:ISOMIP_mass_and_topog}b respectively, with the grounding line position drawn in for reference. The configuration is the same as that of the Ocean0 setup in the MISOMIP, with three changes made:
\begin{enumerate}
\item The `calving criteria' used in the MISOMIP study (which states that all points in the ice shelf with thickness less than 100m are set to zero thickness) has not been used.
\item The ice shelf has been thickened on the flanks of the domain, so that the latitude of the grounding line increases away from the center of the ice shelf.
\item The ice shelf is configured to be symmetric about its meridional center line ($x=\frac{L_{x}}{2}$). This was achieved by using the average of the left and right flanks of the ice-shelf thickness. 
\end{enumerate}
These three changes were made in order to make the circulation beneath the ice shelf easier to interpret.

%\textbf{Initializing Lagrangian elements:}\\
\subsection{Initializing Lagrangian elements:}
The idealized ice shelf is constructed out of Lagrangian ice elements.
Ice elements are set to be hexagonal and are arranged in a regular staggered lattice (as discussed in Section 2.3). The sides of the gridded hexagons are initialized with length $S=0.98$ km. Recall that the element size must be smaller than the ocean grid spacing (i.e.: $2S<\Delta x$ = 2 km). 
Gaps along the boundaries are filled in using smaller elements so that the total ice-shelf area is preserved. The initial mass of the ice elements is determined by a preprocessing inversion step, which is the inverse of the 'mass-spreading' interpolation procedure discussed in Section 2.3.

Figure \ref{fig:ISOMIP_mass_and_topog}c shows what the ice-shelf draft would be if the draft were calculated from the mass of elements in each ocean grid cell without spreading an element's mass across neighboring cells (i.e.: treating elements as point masses). Figure \ref{fig:ISOMIP_mass_and_topog}b shows the draft after spreading the mass across grid cells. When the mass spreading interpolation scheme is not used, grid artifacts seen in the ice-shelf draft (Figure \ref{fig:ISOMIP_mass_and_topog}c). The grid artifacts are much reduced when the mass spreading interpolation is used (Figure \ref{fig:ISOMIP_mass_and_topog}b).

\subsection{Ocean model setup}
The LTAM model is coupled to the MOM6 ocean model.
The ocean model is run using vertical coordinate system which is a hybrid between a sigma-level and a z-level coordinate. In particular, model layers bend underneath the ice shelf as they would in a sigma-coordinate model, but collapse to zero thickness when they intersect with bottom topography, as they would in a z-level model. The coordinate system was achieve using ALE regridding-remapping scheme \citep{White2009}. The model uses a horizontal resolution of 2km, and 72 vertical layers.
All simulation were repeated using the ocean model run layered mode (results were similar and are not presented here).
%(select results are shown in the supplementary material). 
Ocean parameters are as specified in the MISOMIP configuration \citep{Asay-Davis2016}, and are shown in Table 1. The simulation is initially at rest, with horizontally uniform initial ocean temperature and salinity profiles that vary linearly between specified surface and bottom values. The initial surface temperature and salinity are  $T_{t}=-1.9^{\circ}$C and $S_{t}=33.8$ psu, respectively. The initial temperature and salinity at the depth $H_{ocean}$ are $T_{b}=1.0^{\circ}$C and $S_{b}=34.7$ psu. The maximum ocean depth is $H_{ocean}=720$ m.  A sponge layer is used on the northern boundary, which relaxes back to the initial temperature and salinity with a relaxation time scale of $T_{sponge}=0.1$ days. Melting is set to zero for ocean cells where the ocean column thickness is less than 10m to avoid using more energy to melt ice than is present in the water column.

\subsection{Spinup and iceberg calving:}
% \subsection{Tabular iceberg calving}
The model is spun up for 5 years with all ice elements being held stationary. After spinup, a giant tabular iceberg is 'broken off' from the ice shelf, and allowed to drift into the open ocean. This is achieved by allowing all ice elements initially within a 14.4 km radius of the center of the ice front to move freely while the other ice elements continue to be held stationary. Ice elements less than 12 km from the center of the ice front, are bonded together to form a semi-circular tabular iceberg. A ring of elements whose distance, $d$, from the ice front center obeys $12 \textrm{ km} \leq   d \leq14.4 \textrm{ km}$, are allowed to move freely, but have all their bonds removed. Elements in this half annulus represent fragments of the ice shelf which calve into small pieces during a large calving event. Breaking the bonds of these surrounding elements allows the tabular iceberg to move away from the ice-shelf cavity more easily.

A wind stress $\vec{\tau} = < 0.05, 0.05 >$  $\frac{N}{m^{2}}$ is applied to drive the tabular iceberg away from the ice-shelf cavity. Perturbation experiments were performed using other wind stress values.

%\begin{enumerate}
%\item
%\textbf{Tabular iceberg with fixed velocity}:\\
%In this simulation, the velocity of the non-stationary elements is fixed to u=0m/s, v=0.01m/s. This simulation allows us to study the response of the ocean to disturbance caused by a tabular iceberg moving through the water (with an fixed velocity). This simulation is not fully coupled since the velocity of the elements is prescribed.
%\item
%\textbf{Tabular iceberg with bonded elements}:\\
 %All non-stationary ice elements are bonded together by the numerical bonds. This experiments is fully-coupled in that the ice elements are driven the ocean, and the pressure exerted by the ice elements drives ocean flow. In order to make the iceberg drift away from the ice shelf, we imposed a zonal wind stress of $\tau_{x}$ = 0.1 $\frac{N}{m^{2}}$. .
%\item
%\textbf{Tabular iceberg with unbonded elements}:\\
%This simulation is identical to the bonded elements simulation, except that all the numerical bonds are broken, so that the elements are free to move away from one another.
%\end{enumerate}

%\section{Results}
%\subsection{Uncoupled simulation}
%Figure \ref{fig:Uncoupled_sim} shows the results of the uncoupled simulation described in subsection 3.1. The figure shows snapshots of the positions of ice elements at after t=0, 300, 600 and 900 days.  During the simulation, the ice elements drift westward with the prescribed ocean current until the reach the land boundary on the western side of the domain.  Once the elements reach the western coastline, the elements become trapped against the coast and can not move further. When new elements approach the elements at the coast, they interact with the elements already at the coast, and are repulsed. The elements eventually settle in an equilibrium position where there is minimal overlap between elements.
%
%The tabular icebergs also drift westward towards the coast. The elements within the tabular icebergs which arrive first at the coastline  bump up against the coastline (or elements already at the coast). This interaction causes the tabular icebergs to rotate, and change directions. The tabular icebergs continue to drift and interact with other elements until they too settle into an equilibrium position and come to rest. The rotation of the tabular icebergs confirms the this solid body rotation can me modeled using elements and bonds, without having to include an equation for the angular momentum of the elements (as discussed in \citet{Jakobsen2001}).
%
%Comparing the northern and southern halves of the simulation shows how large tabular icebergs can interfere with the trajectories and movement of smaller icebergs (or sea ice flows). These types of interactions will likely play an important role in ice choked fjords [reference] and Antarctic coastal regions [reference].

\section{Model Results}

%\subsection{Mass spreading}
%When the model is run, the average ice draft in an ocean grid cell, is calculated from the total mass of ice elements in that grid cell by assuming that the ice elements are in hydrostatic balance. The total mass is found by summing up over all ice elements located in that grid cell, while allowing elements to spread their mass across neighboring grid cells (as discussed in Section 2.3). Figure \ref{fig:ISOMIP_mass_and_topog}b shows the draft of the ice shelf above the ice shelf cavity in the coupled static ice shelf simulation using this interpolation method, while Figure \ref{fig:ISOMIP_mass_and_topog}c shows the draft which is calculated without spreading an elements mass across neighboring cells (i.e.: treating elements as point masses). When the mass spreading interpolation scheme is not used, grid artifacts seen in the ice shelf draft (Figure \ref{fig:ISOMIP_mass_and_topog}c). The grid artifacts are much reduced when the mass spreading interpolation is used (Figure \ref{fig:ISOMIP_mass_and_topog}b). In simulations without mass spreading, irregularities in ice draft cause irregularities in the ocean surface pressure, which cause spurious motion in the ocean (not shown). This is much reduced or absent when mass spreading is used.

%\subsection{Coupled simulation}

%Figures \ref{fig:Melt_comparison} show the time-averaged melt rates of both the Lagrangian and Eulerian ice shelf models, averaged over the first 5 years of the simulation.   The melt rates of the Eulerian and Lagrangian models are very similar with slight difference showing up at the grounding line and ice front (Figure \ref{fig:Melt_comparison}). The slight difference at the grounding line are a result of the fact that elements in the LTAM model are smaller than a grid cell in length. This means that ice element near the grounding line may be ungrounded in the LTAM model, while the grid cell is considered grounded in the Eularian model. At the ice front, the Eularian model has a straight ice front, while the LTAM model has a staggered ice front, which is a result of the staggered hexagonal packing.

%The small changes in melt at the grounding line and ice front drive slight changes to the circulation beneath the ice shelf, which in turn lead to small difference in the temperature and salinity profiles over time. This is seen by comparing the time-averaged ocean salinity profiles for the Lagrangian and Eularian ice shelves (Figure \ref{fig:Salt_comparison} ) and also  by comparing the the time-averaged barotropic stream function of the two simulations (see Figures \ref{fig:Bt_comparison}) and in the supplementary material).

%\subsection{Decaying ice shelf experiment}
%In the decaying ice shelf experiment, the ice shelf is allowed to decay as it melts. As the ice shelf melts, it decays and the pressure on the top layer of the ocean model decreases (Figure \ref{fig:Decaying_shelf}). The ice shelf decays most rapidly near to the grounding line. Grounding line migration due to melt does not take place since melting is set to zero where the ocean thickness is less that 10m.

%As the ice shelf melts, there is a mass flux into the ocean. The integrated sea surface height north of the ice shelf does not change during the course of the simulation, since the extra mass flux into the ocean, is exactly offset by a decrease mass of the ice shelf, as expected.

%\subsection{Tabular iceberg calving experiment}

During spinup, injection of positive buoyancy due to melting at the base of the ice shelf drives a circulation within the domain (not shown). The circulation compares well with an identical experiment run using an Eularian ice shelf model \citep{Goldberg2012}  (a detailed comparison of the Lagrangian and Eularian ice shelf models is presented in a separate study, and is not shown here).

After spinup, the numerical bonds are severed, and the tabular iceberg and smaller ice fragments calve away from the ice shelf and begin to move (see animations in the supplementary materials). The tabular iceberg drifts to the northward east, driven by the wind and steered by the Coriolis force (Figure \ref{fig:Spread_area}, \ref{fig:Temperature_section_Collapse}). After the iceberg has moved away from the ice shelf, we observed a warming of the surface waters nearby ice front and iceberg edge (Figure \ref{fig:SST_Collapse}). This surface warming is caused by an upwelling of the warmer waters from beneath the ice shelf and iceberg. As the icebergs drifts away from the ice front, these warmer waters remain at the surface, mapping of the wake of the iceberg  (Figure \ref{fig:SST_Collapse}). The signature of upwelling water in the wake of a drifting tabular iceberg bears some similarity to satellite observations of streaks of increased ocean color in the wake of tabular iceberg in the Southern Ocean \citep{Duprat2016}, suggesting that the increased productivity around iceberg may be driven by upwelling water delivering nutrients to the surface.

The motion of the tabular iceberg disturbs the ocean surface, which drives ocean velocities through out the water column (Figure \ref{fig:Velocity_section_Collapse}). The elevated velocities around the tabular iceberg lead to increased vertical mixing in the vicinity of the iceberg, which alters the stratification of the water column (Figure \ref{fig:Temperature_section_Collapse}), heating the upper ocean. The increased ocean velocities and increased surface temperatures cause elevated melt rates at the base of the ice shelf and iceberg (Figure \ref{fig:Collapse_solo_melt}). Large melt rates are observed at the newly calved ice-shelf front and on the rounded side of the iceberg (Figure \ref{fig:Collapse_solo_melt}a), where the iceberg calving has created steep ice cliffs. These sharp ice fronts allow for large ocean currents (Figure \ref{fig:Collapse_solo_melt}c), which drive the elevated melt rates. These elevated melt rates act to smooth out the ice front over time, making the ice cliff less steep. While this is likely a real phenomena that could be observed in nature, we should be wary of the modeled velocities at the ice cliffs, since large changes in ice thicknesses are often associated with numerical pressure gradient errors which can drive spurious motion.

As mentioned above, the direction (and speed) of the iceberg drift is largely determined by the wind speed and direction. Perturbation experiments using different wind stresses showed that for sufficiently large winds, the iceberg drifts to the north east when the zonal component of the wind stresses is positive, and to the north west when the zonal component of the wind stress is negative. For a purely zonal wind stress with $\tau_{x} \leq 0.01 \frac{\textrm{N}}{\textrm{m}^{2}}$, the iceberg does not move away from the ice shelf. 
When the wind is purely offshore ($\tau_{x} = 0.0 \frac{\textrm{N}}{\textrm{m}^{2}}$), a meridional wind stress $\tau_{y} \geq 0.05 \frac{\textrm{N}}{\textrm{m}^{2}}$ is needed to move the tabular iceberg away from the ice shelf.
While this result is partly an artifact of the artificial shape of the calving iceberg, it is also consistent with \citet{Bassis2013} who noted that calving is a two step process consisting of (i) ice-shelf breaking and (ii) iceberg detachment. The results here suggest that strong (cross-shore) winds may be required to drive large tabular icebergs away from their mother glaciers.

Finally, we note that the numerical bonds in the LTAM model are needed in order to allow the tabular iceberg to retain its shape. Comparing the iceberg calving simulation with an identical simulation where all numerical bonds have been removed, shows that in the absence of the bonds, the ice elements quickly disperse (Figure \ref{fig:Bond_vs_No_bond}). In this case, the model behavior is more similar to an ice-shelf disintegration and does not create a cohesive tabular iceberg.

%\subsubsection{Tabular iceberg with fixed velocity}
%\textbf{Tabular iceberg with fixed velocity}\\
%The fixed iceberg velocity calving experiment demonstrates that the model is able to simulate a tabular iceberg being calved from an ice shelf.  After the iceberg calves, it moves northwards away from the ice shelf with a prescribed velocity. After the iceberg has calved away from the ice shelf, we observed a warming of the surface waters near the near the ice front and iceberg edge. This surface warming is caused by an upwelling of the warmer waters from beneath the ice shelf and iceberg. As the icebergs drifts away from the ice front, these warmer waters remain in the wake of the iceberg (Figures \ref{fig:Calving_fixed_top}). The signature of upwelling water in the wake of a drifting tabular iceberg bears some similarity to satellite observations of streaks of increased ocean color near to tabular iceberg in the Southern Ocean [reference],  suggesting that the increased productivity around iceberg may be driven by upwelling water delivering nutrients to the surface. The motion of the iceberg also drives barotropic velocities through out the water column, which disturb the initial stratification of the water column (Figures \ref{fig:Calving_fixed_side}).

%\subsubsection{Fully coupled tabular iceberg}
%\textbf{Fully coupled tabular iceberg}\\
%The necessity of bonds between ice elements is demonstrated by comparing two simulations where an iceberg calves, and is allowed to evolve freely with the flow. In the Bonded simulation, the elements in the calved iceberg are bonded together, while in the Unbonded simulation the all bonds between elements in the tabular iceberg are broken.

%The motions of elements in the tabular iceberg in these two simulations are rather different. 
%In the Bonded simulation, the ice element in the tabular iceberg remain close to one another, and the tabular iceberg retains its shape (Figure \ref{fig:Bond_vs_No_bond}a,c). The elements are forced by the melt-water-driven current at the ice front, which moves towards the north-west (see Figure \ref{fig:static_solo} in the supplementary material). This current, together with the Coriolis force, cause the iceberg to rotate before moving away from the ice shelf (Figure \ref{fig:Bond_vs_No_bond}a).

%In the Unbonded simulation, the ice elements in tabular iceberg are not constrained to move together. Instead, the elements spread out and drift towards the north-west for the same reason discussed above. Since the ice elements are not closely, packed together, they apply a smaller pressure to the ocean surface, spread over a larger area (Figure \ref{fig:Bond_vs_No_bond}c). The type of behavior could be relevant in simulating ice shelf disintegrations, rather than tabular iceberg calving.

 \section{Summary}
In study above we have presented a novel method that can be used to representing tabular icebergs in numerical ocean models. In this framework, tabular icebergs (or ice shelves) are constructed out of Lagrangian particles which are held together by numerical bonds. In this way icebergs are not treated as point particles (as in previous iceberg models), but rather are explicitly resolved in the ocean. This allows the icebergs to apply pressure to the ocean surface over a wide area (larger than a grid cell), and affect circulation patterns in a way that is more physically realistic.

A demonstration of the tabular iceberg model was performed using an idealized semi-circular iceberg drifting away from an idealized ice shelf. The results showed that explicitly resolving the iceberg in the ocean allowed for a complex interaction between the tabular iceberg and the surrounding ocean. In our setup, the tabular iceberg is driven way from the ice shelf by ocean currents, wind stress, and the Coriolis force. As the iceberg moves through the water, it disturbs the ocean surface, driving barotropic ocean currents. The motion of the iceberg and melt beneath the iceberg also drive upwelling along the sides of the iceberg, which entrains ambient water and causes a warming of the surface ocean in the wake of the iceberg. The changing ocean conditions feed back onto the iceberg, affecting its motion and melt rates. The highest melt rates are observed at edge of the iceberg which has the steepest ice cliff. These have the effect of smoothing out the ice edge over time.
Simulations without using numerical bonds showed that these bonds are essential for allowing the iceberg to move as a unit. We also saw that by breaking these numerical bonds we can simulate iceberg fracture, which is important process which rapidly increases the rate of iceberg decay.

The model presented here is potentially a step towards including tabular icebergs in global general climate models (GCM's). However, before this can be done, there are a number of other outstanding questions to be resolved. Firstly, it is currently unclear how and when one would introduce tabular icebergs into climate simulations, especially if the GCM does not use an active ice shelf from which the tabular icebergs can calve. For simulations with ice shelves, a calving law would be needed to decide when to release the tabular iceberg into the ocean. The question of what calving law to use is a topic of ongoing research \citep{Benn2007, Alley2008, Levermann2012, Bassis2013} and is still unresolved. One potential way bypass this problem would be to run hindcast simulations using historically observed calving events. Another question is how and when to break the bonds within the freely floating icebergs to simulation iceberg breakup. Without a rule for iceberg breakup, the tabular icebergs would likely drift to unrealistically low latitudes. There are also a number of outstanding questions involving the interaction between tabular icebergs and sea ice, ocean biogeochemistry and iceberg grounding. None-the-less, the technical framework described in this article is potentially a using step towards including tabular icebergs in global general climate models (GCM's).

\section{Appendix A}
\subsection{Environmental forces on ice elements}
The non-interactive forces on an ice element are as described in \citep{Martin2010}, and are repeated here for completeness. 
The forces on an element due to air (a), ocean (o) and sea ice (si) drag are given by
\begin{equation}
(\vec{F}_{a}) = \rho_{a} (0.5 c_{a,v} W F + c_{a,h} L W) | \vec{u}_{a} - \vec{u}|(\vec{u}_{a} - \vec{u}),
\end{equation}
\begin{equation}
(\vec{F}_{o}) = \rho_{o} (0.5 c_{o,v} W (D-T_{si})F + c_{o,h} L W) | \vec{u}_{o} - \vec{u}|(\vec{u}_{o} - \vec{u}),
\label{eq:ocean_drag_eq}
\end{equation}
\begin{equation}
(\vec{F}_{si}) = \rho_{si} (0.5 c_{si,v} W T_{si}F + c_{si,h} L W) | \vec{u}_{si} - \vec{u}|(\vec{u}_{si} - \vec{u}).
\end{equation}
Here $\rho_{a}$, $\rho_{o}$, $\rho_{si}$, are the density of air, ocean and sea ice, respectively. $c_{a,v}$, $c_{o,v}$ and $c_{si,v}$ are the vertical drag coefficients with air, ocean and sea ice, while $c_{a,h}$, $c_{o,h}$ and $c_{si,h}$ are the respective horizontal drag coefficients. $\vec{u}_{a}$, $\vec{u}_{o}$, $\vec{u}_{si}$, are the velocities air, ocean and sea ice, respectively. L, W, T, F and D are the length, width, thickness, freeboard and draft of the ice element. The element thickness is related to the draft and freeboard by $T=F+D$ and $D=\frac{\rho}{\rho_{o}} T$, where $\rho$ is the ice element density.

The  wave radiation force $(\vec{F}_{R})$ is given by
\begin{equation}
\vec{F}_{R} = \frac{1}{2}\rho_{o} c_{r} g a \textrm{ min}(a,F)  \textrm{ } 2\frac{WL}{W+L} \frac{\vec{v}_{a}}{|\vec{v}_{a} |}
\end{equation}
where g is the acceleration due to gravity, a is the wave amplitude empirically related to the wind speed by  $a=0.010125|\vec{v}_{a} -  \vec{v}_{o}| $, and $c_{wd}$ is the wave drag coefficient defined as 
\begin{equation}
c_{wd}=0.06 \textrm{ min} \Bigg( \textrm{ max} \Bigg[0,\frac{L-L_{c}}{L_{t}-L_{c} } \Bigg]  , 1    \Bigg), 
\end{equation}
where $L_{w}=0.32|\vec{v}_{a} -  \vec{v}_{o}|^{2}$ is an empirical wave length, $L_{c}=0.125 L_{w}$ is the cutoff length, and  $L_{t}=0.25 L_{w}$ is the upper limit.

The pressure gradient force is approximated as a force due to sea surface slope and given by
\begin{equation}
\vec{F}_{SS} = -M g \vec{\nabla} \eta
   \end{equation}
where $\eta$ is the sea surface height.

\subsection{Melt rate parametrization}
As discussed in Section 2.5, unbounded ice elements in the LTAM model decay according to parameterizations for iceberg decay typically used in iceberg drift models \citep{Martin2010}, while ice elements within larger ice structures have only a basal melt given by the three equation model \citep{Holland1999} .

For unbonded ice elements, the element thickness decays due to basal melt at a rate $M_{b}$, while the length and width of the elements decay as a result of melt due to wave erosion, $M_{e}$, and melt due to buoyant convection, $M_{v}$. Following \citet{Gladstone2001} and \citet{Martin2010}, the basal melt rate, wave erosion melt rate, and buoyant convection melt rate are parameterized by 
\begin{equation}
M_{b} = 0.58 | \vec{v} - \vec{v}_{0}|^{0.8} \frac{\tilde{T}_{0} - \tilde{T}}{L^{0.2}}
\end{equation}
\begin{equation}
M_{e} = \frac{1}{12} S_{s} \Bigg(  1 + \cos{[\pi A_{i}^{3}]}  \Bigg) \Bigg( \tilde{T}_{0} +2 \Bigg ),
\end{equation}
\begin{equation}
M_{v} =  \Bigg(7.62 \times 10^{-3}  \Bigg )   \tilde{T}_{0} + \Bigg( 1.29 \times 10^{-3} \Bigg)\tilde{T}_{0}^{2}.
\end{equation}
$\tilde{T}$ is the effective iceberg temperature and is set to $\tilde{T}=4^{\circ}$C, $\tilde{T}_{0}$ is the temperature at the top of the ocean,  $A_{i}$ is the sea ice area fraction, and $S_{s}$ is the sea state, which is given by the Beaufort scale
\begin{equation}
S_{s} =  \frac{2}{3}  | \vec{u}_{a} - \vec{u}_{o} |^{\frac{1}{2}}  +\frac{1}{10}   | \vec{u}_{a} -\vec{u}_{o} |
\end{equation}
All three melt rates are in units of meters per day.

For elements inside larger structures, the melt due to wave erosion  and melt due to buoyant convection are set to zero, and the basal melt, $M_{s}$ is given by the standard three equation model \citep{Holland1999}. 
%\begin{equation}
%M_{s} =\frac{ 1}{L_{ice} \rho_{ice} }    \rho_{o} c_{po}\gamma_{T}  \Bigg(  T_{o} -T_{f}(S_{o},p_{o})  \Bigg)
%\end{equation}
%where $T_{o}$, $S_{o}$, $p_{o}$ are the ocean mixed layer temperature, salinity and pressure.  $T_{f}(S_{ml}, p_{ml})$ is the salinity and pressure dependent freezing point of sea water at salinity $S_{o}$ and pressure $p_{o}$. $\gamma_{T}$ is the turbulent transfer coefficient given by
%\begin{equation}
%\gamma_{T}=\frac{1}{\Gamma_{Turb} +\Gamma^{T}_{Molec}}
%\end{equation}

\section{Appendix B}
\subsection{Modified Verlet Algorithm}
The LTAM model uses a version velocity verlet time-stepping algorithm, which has been modified to allow part of the forcing to be calculated implicitly. The traditional velocity verlet algorithm is commonly used in molecular dynamics, as it is simple to implement, second order accurate and computationally efficient \citep{Swope1982, Omelyan2002}.
Here we modify the tradition scheme to allow for the drag forces to be modeled implicitly, which prevents large accelerations for element's whose mass approaches zero. 
To do this, we modify the traditional verlet algorithm to include both an implicit and explicit acceleration, $a=a^{exp} +a^{imp}$. The explicit acceleration, $a^{exp}$ includes all forcing terms which depend only on the previous time step, while the implicit acceleration, $a^{imp}$  includes forcing terms which depend on the current time step (in particular the drag and Coriolis forces).

%1) $ x(t_{n+1}) = x(t_{n}) + u(t_{n}) \Delta t  + \frac{1}{2} \Delta t^{2} a(t_{n}) $.\\
%2) Calculate $a(t_{n+1})$\\
%3) $u(t_{n+1}) = u(t_{n}) + \frac{\Delta t}{2} \Bigg(a(t_{n}) + a(t_{n+1}) \Bigg)$\\
%Note that in step 2, $a(t_{n+1})$ is an explicit function of  $x(t_{n+1})$ and other quantites evaluated at time $t_{n}$.

Using a time step of $\Delta t$, and subscripts to denote the time step (so that $t_{n+1}= t_{n} + \Delta t$), the modified velocity verlet scheme can be written as:\\
1) $ x_{n+1} = x_{n} + u_{n} \Delta t  + \frac{1}{2} \Delta t^{2} \Bigg(a^{exp}_{n} + a^{imp}_{n} \Bigg) $.\\
2) Calculate $a^{exp}_{n+1}$, which is a function of $x_{n+1}$, and other quantities measured at time $t_{n}$. \\
3) Calculate $a^{imp}_{n+1}$ and $u_{n+1} = u_{n} + \frac{\Delta t}{2} \Bigg(a^{exp}_{n} + a^{exp}_{n+1} \Bigg) +  (\Delta t) a^{imp}_{n+1}$\\
This scheme reduces to the traditional velocity verlet when $a^{imp}$ is set to zero. Note that $a^{exp}_{n+1} = a^{exp}_{n+1}(x_{n+1}, t_{n})$ is an explicit function of  $x_{n+1}$ and other quantities evaluated at time $t_{n}$, while $a^{imp}_{n+1}=a^{imp}_{n+1}(u_{n+1}, x_{n+1}, t_{n})$ additionally depends on $u_{n+1}$, and needs to be solved implicitly. For this reason, in step 3 $a^{imp}_{n+1}$ and $u_{n+1}$ need to be solved simultaneously, as described in the next subsection.

In equation (1), the forces due to ocean drag, atmospheric drag and sea ice drag are treated implicitly. The force due to sea surface slope and wave radiation are treated explicitly. The Coriolis term is handled using Crank-Nicolson scheme so that half of the effect is implicit and half is explicit. The elastic part of the interactive forces is treated explicitly, while the interactive damping is handled semi-implicitly in that the drag force on element A by element B depends on the velocities of elements A and B evaluated at time $t_{n+1}$ and $t_{n}$, respectively.

 %A full description of how the scheme is implemented is provided below.

 \subsection{Solving for the velocity implicitly}
Since this modified scheme contains some forcing terms which are handled implicitly, steps 2 and 3 need to be solved together.  We demonstrate how this is done, using a simplified one-dimensional version of equation (1), neglecting the atmospheric drag, sea ice drag and Coriolis force, so that the only implicitly treated term is the ocean drag. In this demonstration, we use a superscript to denote the ocean drag force, $F^{o}$ and ocean velocity, $u^{o}$ to avoid confusion with the subscripts indicating time step. We also define an explicit force $F^{exp}$, which accounts for all forces not proportional the element velocity. With these simplifications, the implicit and explicit accelerations are
 \begin{equation}
a^{exp} =\frac{1}{M} ( \vec{F}^{exp} )
\end{equation}
 \begin{equation}
a^{imp} =\frac{1}{M} (F^{o})
\end{equation}
The ocean drag force at time $t_{n+1}$ is modeled (mostly) implicitly as 
\begin{equation}
\label{eq:Drag}
F^{o}_{n+1} = \tilde{c}^{o} |u^{o}_{n} -u_{n} | (u^{o}_{n} -u_{n+1} ),
\end{equation}
where $\tilde{c}^{o}$ is the effective drag coefficient, accounting for the dimensions of the ice element (see equation \ref{eq:ocean_drag_eq}).

%The consequences of the drag not being fully implicit are discussed below. 
Step 3 of the modified velocity verlet scheme can be rewritten by introducing an intermediate velocity $u^{*}$, which only depends on the velocity and acceleration at time $t_{n}$,
 \begin{equation}
 u^{*}_{n} =  u_{n}  + \frac{1}{2} (\Delta t) a_{n}. 
 \end{equation}
Using this, the updated velocity (Step 3) can be written 
 %\begin{equation}
 %x(t_{n+1}) = x(t_{n}) + v^{*}(t_{n}) \Delta t 
 % \end{equation}
 \begin{equation}
u_{n+1} = u^{*}_{n} + \frac{\Delta t}{2} a^{exp}_{n+1}+  (\Delta t) a^{imp}_{n+1}.
 \end{equation}

Including the forcing terms into this equations gives
 \begin{equation}
u_{n+1} = u^{*}_{n} + \frac{\Delta t}{2M} ( F^{exp} _{n+1}) +   \frac{\Delta t}{M}     \Bigg(  c_{w} |u^{o}_{n} -u_{n} | (u^{o}_{n} -u_{n+1} )   \Bigg)
 \end{equation}
Solving for $u(t_{n+1})$ in terms of quantities which only depend on the previous time step gives
 \begin{equation}
u_{n+1} = \frac{u^{*}_{n} + \frac{\Delta t}{2M} ( F^{exp}_{n+1}) +   \frac{\Delta t}{M}    \Bigg(  c_{w} |u^{o}_{n} -u_{n} | (u^{o}_{n} )   \Bigg)}{  \Bigg( 1+  \frac{\Delta t}{M}       c_{w} |u^{o}_{n} -u_{n} |    \Bigg)}
 \end{equation}
Once the $u_{n+1}$  has been found, it can be used to calculated the explicit and implicit accelerations, which are required for the next time step.

Finally, we note that the the drag term (equation \ref{eq:Drag}) is not entirely implicit, since the element velocity inside the absolute value is evaluated at time $t_{n}$, rather than at time $t_{n+1}$. This is done so that we can solve for the updated velocity analytically. One consequence of this is that it gives rise to a small oscillation in the element velocity. This oscillation is addressed by using a predictive corrective scheme: once you have solved for a first guess of the velocity at time $t_{n+1}$, this estimate of the velocity is used to update the estimate of the drag force (i.e.: inside the absolute value signs). Using the updated drag, you can now repeat the process described above to find an improved estimate of the velocity. We found that two iterations were sufficient to remove the unwanted oscillation.

The procedure described in this section is easily extended to include more forcing terms and two dimensions (where it involves inverting a $2 \times 2$ matrix).

\section{Appendix C}
\textbf{Connecting bonds across processor boundaries}\\
Since the LTAM model can be parallelized across multiple processors, it often happens that two elements on different processes are bonded together. Keeping track of numerical bonds across processor boundaries requires a lot of book keeping. In this section we describe the how LTAM model avoids segmentation faults by severing and reconnecting bonds at the edge of a processor during halo updates.

%When a bonded element moves off the edge of a processor, the bond needs to be broken (to avoid segmentation faults). Before the next time step, the code performs a `halo update', and a copy of the element is written to the halo of the original processor. After this, the bond which was broken at the previous time step, needs to be reconnected. This processes of severing and reconnecting bonds requires a lot of bookkeeping, some of which are described below.

The basics of the bond bookkeeping work as follows:
consider an element A and an element B that are bonded together. Each element has a copy of the bond (a piece of memory which describes the bond between the two elements), which is stored with the element. Let A-B be the bond stored by element A, and B-A be the bond stored by element B. Bond A-B contains a pointer which points to element B and bond B-A contains a pointer which points to element A.

Consider a situation where element A and B are originally on Processor 1, and then element B moves to Processor 2. When this occurs, the memory assigned to element B on processor 1 is removed, and is allocated on Processor 2. This means that the pointer to element B in bond A-B (stored in element A on Processor 1) is no longer assigned. Similarly, the pointer to element A in bond B-A (stored in element B on Processor 2) is no longer assigned. Before the next time step, a halo update occurs, so that the there is a copy of element A in the halo of Processor 2 and a copy of element B in the halo of Processor 1. After the halo update, the bonds A-B and B-A have to be reconnected on both Processor 1 and 2. To aid in reconnecting the bonds, a copy of the grid cell number of element B is stored in the bond A-B and a copy of the grid cell number of element A is stored in the bond B-A. We refer to this as the `most recent address'. Before a bond is moved from one processor to another, the `most recent address' is updated, so that the bond can be reconnected later. To reconnect bond A-B on Processor 1 (for example), we find the most recent address of element B, and search through the list of elements in the grid cell corresponding to the most recent address of element B until element B is found. The pointer to element B in bond A-B is reassigned and the bond is said to be connected.

Once all bonds are reconnected, a bond quality control is done where we loop through all bonds and check that they are working properly. To check that a bond is working properly is a four step process. For example, consider the bond A-B stored in element A on Processor 1. To check the quality of this bond, we use the following four steps:
\begin{enumerate}
\item Check that the pointer to element B is assigned on bond A-B (stored on element A).
\item Check that the corresponding bond B-A exists on element B.
\item Check that a pointer to element A exists in this bond B-A.
\item Check that the element A which is being pointed to is the same element A where you started.
\end{enumerate}
All four of these tests must pass in order to make sure that the memory is correctly assigned. A useful tool in this process is that each element is assigned a unique number so that elements are easily identified.

\clearpage

\section{Possibly add}
\subsection{Orientation}
Possibly add a section where we spread mass using the orientation of the icebergs.

\newpage
\begin{thebibliography}{9}
\expandafter\ifx\csname natexlab\endcsname\relax
  \def\natexlab#1{#1}\fi
\expandafter\ifx\csname selectlanguage\endcsname\relax
  \def\selectlanguage#1{\relax}\fi
%
\bibitem[Asay-Davis et al (2016)]{Asay-Davis2016}
Asay-Davis, X. S., S. L. Cornford, B. K. Galton-Fenzi, R. M. Gladstone, G. H. Gudmundsson, D. M. Holland, P. R. Holland, and D. F. Martin (2016), Experimental design for three interrelated marine ice sheet and ocean model intercomparison projects: MISMIP v. 3 (MISMIP+), ISOMIP v. 2 (ISOMIP+) and MISOMIP v. 1 (MISOMIP1). {\it Geoscientific Model Development} 9, no. 7: 2471.
%
\bibitem[Arrigo et al (2002)]{Arrigo2002}
Arrigo, K. R., G. L. van Dijken,  D. G. Ainley, M. A. Fahnestock, and T. Markus (2002). Ecological impact of a large Antarctic iceberg. {\it Geophys. Res. Let.}, 29(7).

\bibitem[Alley et al (2008)]{Alley2008}
Alley, R. B., H. J. Horgan, I. Joughin, K. M. Cuffey, T. K. Dupont, B. R. Parizek, S. Anandakrishnan, and J. Bassis (2008), A simple law for ice-shelf calving. {\it Science} 322, no. 5906, 1344-1344.

\bibitem[Bassis and Jacobs (2013)]{Bassis2013}
Bassis, J. N., and S. Jacobs (2013), Diverse calving patterns linked to glacier geometry. {\it Nature Geoscience}, 6(10), 833-836.

\bibitem[Benn et all (2007)]{Benn2007}
Benn, D. I., C. R. Warren, and R. H. Mottram (2007). Calving processes and the dynamics of calving glaciers. {\it Earth-Science Reviews}, 82(3), 143-179.

\bibitem[Bigg et al (1997)]{Bigg1997}
Bigg, G. R., Wadley, M. R., Stevens, D. P., and Johnson, J. A. (1997), Modeling the dynamics and thermodynamics of icebergs. {\it Cold Regions Science and Technology}, 26(2), 113-135.

\bibitem[Borstad et al (2012)]{Borstad2012}
Borstad, C. P., A. Khazendar, E. Larour, M. Morlighem, E. Rignot, M. P. Schodlok, and H. Seroussi (2012), A damage mechanics assessment of the Larsen B ice shelf prior to collapse: Toward a physically-based calving law, {\it Geophys. Res. Lett.}, 39, L18502

\bibitem[Biddle et al (2015)]{Biddle2015}
Biddle, L. C., J. Kaiser, K. J. Heywood, A. F. Thompson and A. Jenkins (2015), Ocean glider observations of iceberg-enhanced biological productivity in the northwestern Weddell Sea, {\it Geophys. Res. Lett.}, 42, 459–465.

\bibitem[De Rydt and Gudmundsson (2016)]{De_Rydt2006}
De Rydt, J., and G. H. Gudmundsson (2016), Coupled ice shelf ocean modeling and complex grounding line retreat from a seabed ridge. {\it J. of Geophys. Res.: Earth Surface}, 121(5), 865-880.
%
\bibitem[Dunne et al (2012)]{Dunne2012}
Dunne, J.P., J.G. John,, A.J. Adcroft, S.M. Griffies, R.W. Hallberg, E. Shevliakova, R.J. Stouffer, W. Cooke, K.A. Dunne, M.J Harrison, and J.P. Krasting (2012), GFDL's ESM2 global coupled climate-carbon Earth System Models. Part I: Physical formulation and baseline simulation characteristics. {\it J. of Climate}, 25(19), 6646-6665.
%
\bibitem[Depoorter et al (2013)]{Depoorter2013}
Depoorter, M. A., J. L. Bamber, J. A. Griggs, J. T. M. Lenaerts, Stefan RM Ligtenberg, M. R. van den Broke, and G. Moholdt (2013), Calving fluxes and basal melt rates of Antarctic ice shelves. {\it Nature}, 502(7469), 89-92.

\bibitem[Determan and Gerdes (1994)]{Determan1994}
Determan J.,  Gerdes R. (1994), Melting and freezing beneath ice shelves: implications from a three-dimensional ocean-circulation model. {\it  Ann. Glaciol.}, 20, 413-419.

\bibitem[Dowdeswell and Bamber (2007)]{Dowdeswell2007}
Dowdeswell, J. A., and J. L. Bamber (2007), Keel depths of modern Antarctic icebergs and implications for sea-floor scouring in the geological record. {\it Marine Geology}, 243(1), 120-131.

\bibitem[Duprat et al (2016)]{Duprat2016}
Duprat, L. P., G. R. Bigg, and D. J. Wilton (2016), Enhanced Southern Ocean marine productivity due to fertilization by giant icebergs. {\it Nature Geoscience.}

\bibitem[Eckert (1950)]{Eckert1950}
Eckert, E. R. G. (1950). Introduction to the Transfer of Heat and Mass. McGraw-Hill.

\bibitem[Gladstone et al (2001)]{Gladstone2001}
Gladstone, R. M., G. R. Bigg, and K. W. Nicholls. (2001), Iceberg trajectory modeling and meltwater injection in the Southern Ocean (1978–2012). {\it J. of Geophys. Res.: Oceans}, 106(C9), 19903-19915.

\bibitem[Goldberg et al (2012)]{Goldberg2012}
Goldberg, D. N., C. M. Little, O. V. Sergienko, A. Gnanadesikan, R. Hallberg, and M. Oppenheimer (2012), Investigation of land ice?ocean interaction with a fully coupled ice-ocean model: 1. Model description and behavior. {\it J. of Geophys. Res.: Earth Surface}, 117(F2).
%
\bibitem[Gladish et al (2012)]{Gladish2012}
Gladish, C. V., D. M. Holland, P. R. Holland, and S. F. Price (2012), Ice-shelf basal channels in a coupled ice/ocean model. {\it J. of Glaciol.}, 58(212), 1227-1244.

\bibitem[Grosfeld et al (1997)]{Grosfeld1997}
Grosfeld K., R. Gerdes, J. Determan (1997), Thermohaline circulation and interaction between ice shelf cavities and the adjacent open ocean. {\it J. Phys. Oceanogr.,  \textbf{102}, C7, 15959-15610}.

\bibitem[Grosfeld and Sandhäger (2004)]{Grosfeld2004}
Grosfeld, K., and H. Sandhäger, (2004). The evolution of a coupled ice shelf–ocean system under different climate states. {\it Global and Planetary Change}, 42(1), 107-132.

\bibitem[Hallberg et al (2013)]{Hallberg2013}
Hallberg, R., A. Adcroft, J. P.  Dunne, J. P., Krasting, R. J., and Stouffer (2013), Sensitivity of twenty-first-century global-mean steric sea level rise to ocean model formulation. {J. of Climate}, 26(9), 2947-2956.

\bibitem[Holland and Jenkins (2001)]{Holland2001}
Holland D. M.,  Jenkins A. (2001), Adaptation of an isopycnic coordinate ocean model for the study of circulation beneath ice shelves. {\it Mon. Wea. Rev}., 129, 1905-1927.

\bibitem[Holland and Feltham (2006)]{Holland2006}
Holland P. R. and  D. L. Feltham (2006), The effects of rotation and ice shelf topography on frazil-laden Ice Shelf Water plumes. {\it  J. Phys. Oceanogr.,}  36, 2312-2327.

\bibitem[Holland and Jenkins (1999)]{Holland1999}
Holland, D. M., and A. Jenkins (1999), Modeling thermodynamic ice-ocean interactions at the base of an ice shelf. {J. of Phys. Oceanogr.} 29.8, 1787-1800.

\bibitem[Hellmer and Olbers (1989)]{Hellmer1989}
Hellmer H.H.,  Olbers D. J. (1989), A two-dimensional model for the thermohaline circulation under an ice shelf. {\it  Antarctic Science},  1, 325- 336.

\bibitem[Henderson and Loe (2016)]{Henderson2016}
Henderson J., J. S. P. Loe (2016), The Prospects andChallenges for Arctic Oil Development. {\it Oil, Gas and Energy Law Journal (OGEL)}, 14 (2)

\bibitem[Hopkins (1996)]{Hopkins1996}
Hopkins, M. A. (1996). On the mesoscale interaction of lead ice and floes. {\it J. of Geophys. Res.: Oceans}, 101(C8), 18315-18326.

\bibitem[Gaskill and Rochester (1984)]{Gaskill1984}
Gaskill, H. S., and J. Rochester (1984). A new technique for iceberg drift prediction. {\it Cold Reg. Sci. Technol.}, 8(3), 223-234.

\bibitem[Jakobsen (2001)]{Jakobsen2001}
Jakobsen, T. (2001). Advanced character physics. {\it In Game Developers Conference}, Vol. 3.

\bibitem[Jenkins et al (2010)]{Jenkins2010}
Jenkins, A., P. Dutrieux, S. S. Jacobs, S. D. McPhail, J. R. Perrett, A. T. Webb, and D. White (2010), Observations beneath Pine Island Glacier in West Antarctica and implications for its retreat. {it Nature Geo.}, 3(7), 468-472.

\bibitem[Jacobs et al (2011)]{Jacobs2011}
Jacobs, S. S., A. Jenkins, C. F. Giulivi, and P. Dutrieux (2011). Stronger ocean circulation and increased melting under Pine Island Glacier ice shelf. {Nature Geo.}, 4(8), 519-523.

\bibitem[Jongma et al (2009)]{Jongma2009}
Jongma, J. I., E. Driesschaert, T. Fichefet, H. Goosse, and H. Renssen (2009), The effect of dynamic-thermodynamic icebergs on the Southern Ocean climate in a three-dimensional model, {\it Ocean Modell.}, 26, 104–113.

\bibitem[Kubat et al (2005)]{Kubat2005}
Kubat I., M. Sayed, S. Savage, T. Carrieres (2005), An operational model of iceberg drift {\it Int. J. Off. Polar Eng.}, 15 (2), 125–131

\bibitem[Lewis and Perkin (1986)]{Lewis1986}
Lewis E.L. and R.G. Perkin (1986), Ice pumps and their rates. {\it J. of Geophys. Res.}, 91, 11756-11762.

\bibitem[Losch (2008)]{Losch2008}
Losch, M. (2008). Modeling ice shelf cavities in az coordinate ocean general circulation model. {\it J. of Geophys. Res.: Oceans}, 113(C8).

\bibitem[Li et al (2014)]{Li2014}
Li, B., H. Li, Y. Liu, A. Wang and S. Ji (2014), A modified discrete element model for sea ice dynamics. {\it Acta Oceanologica Sinica}, 33(1), 56-63.

\bibitem[Liu and Liu (2010)]{Liu2010}
Liu, M. B. and G. R. Liu (2010), Smoothed particle hydrodynamics (SPH): an overview and recent developments. {\it Archives of computational methods in engineering}, 17(1), 25-76.

\bibitem[Lichey and Hellmer (2001)]{Lichey2001}
Lichey, C., and H. H. Hellmer (2001). Modeling giant-iceberg drift under the influence of sea ice in the Weddell Sea, Antarctica. {\it J. of Glaciol.}, 47(158), 452-460.

\bibitem[Levermann et al (2012)]{Levermann2012}
Levermann, A., T. Albrecht, R. Winkelmann, M. A. Martin, M. Haseloff, and I. Joughin. (2012), Kinematic first-order calving law implies potential for abrupt ice-shelf retreat. {\it The Cryosphere}, 6(2), 273-286.

\bibitem[Mountain (1980)]{Mountain1980}
Mountain, D. G. (1980). On predicting iceberg drift. {\it Cold Reg. Sci. Technol.}, 1(3-4), 273-282.

\bibitem[Martin and Adcroft (2010)]{Martin2010}
Martin, T., and Adcroft, A. (2010), Parameterizing the fresh-water flux from land ice to ocean with interactive icebergs in a coupled climate model. {\it Ocean Modelling}, 34(3), 111-124.
%
\bibitem[Marsh et al (2015)]{Marsh2015}
Marsh, R., V. O. Ivchenko, N. Skliris, S. Alderson, G. R. Bigg, G. Madec, A. T. Blaker  Y. Aksenov, B. Sinha, A.C. Coward, and J.L. Sommer (2015), NEMO–ICB (v1. 0): interactive icebergs in the NEMO ocean model globally configured at eddy-permitting resolution. {\it Geoscientific Model Development} 8, no. 5 (2015): 1547-1562.

\bibitem[MacAyeal (1984)]{MacAyeal1984}
MacAyeal D.R. (1984), Thermohaline Circulation Below the Ross Ice Shelf: A Consequence of Tidally Induced Vertical Mixing and Basal Melting. {\it J. Geophys. Res.}, 89, 597-606 
%
\bibitem[Nicholls (1996)]{Nicholls1996}
Nicholls K.W. (1996), Temperature variability beneath Ronne Ice Shelf, Antarctica, from thermistor cables. {\it J. Phys. Oceanogr.},  11, 1199-1210.
%
\bibitem[Nicholls et al (2009)]{Nicholls2009}
Nicholls KW, {\O}sterhus S, Makinson K (2009), Ice-Ocean processes over the continental shelf of the southern Weddell Sea, Antarctica: a review. {\it Rev. Geophys.} 47(3).

\bibitem[Omelyan et al (2002)]{Omelyan2002}
Omelyan, I. P., M. I. Mryglod, and R. Folk (2002), Optimized Verlet-like algorithms for molecular dynamics simulations. {\it Physical Review E}, 65(5), 056706.

\bibitem[Rignot et al (2013)]{Rignot2013}
Rignot, E., S. Jacobs, J. Mouginot, and B. Scheuchl (2013), Ice-shelf melting around Antarctica. {\it Science}, 341, no. 6143 (2013): 266-270.

\bibitem[Robinson et al (2012)]{Robinson2012}
Robinson, N. J., M. J. M. Williams, P. J. Barrett, and A. R. Pyne (2010), Observations of flow and ice-ocean interaction beneath the McMurdo
Ice Shelf, Antarctica, {\it J. Geophys. Res.}, 115, C03025

\bibitem[Pizzolato et al (2012)]{Pizzolato2014}
Pizzolato,  L., S. E. Howell, C. Derksen, J. Dawson, L. Copland (2014), Changing sea ice conditions and marine transportation activity in Canadian Arctic waters between 1990 and  2012, {\it Climatic  Change}  123  (2),  161173.

\bibitem[Pan et al (2013)]{Pan2013}
Pan, W.,  A. M. Tartakovsky, and J. J. Monaghan (2013). Smoothed particle hydrodynamics non-Newtonian model for ice-sheet and ice-shelf dynamics. {\it J. of Comp. Phys.}, 242, 828-842.

\bibitem[Pralong and Funk (2005)]{Pralong2005}
Pralong, A., and M. Funk (2005), Dynamic damage model of crevasse opening and application to glacier calving, {\it J. Geophys. Res.}, 110, B01309.

\bibitem[Sergienko (2013)]{Sergienko2013}
Sergienko, O. V. (2013). Basal channels on ice shelves. {\it J. of Geophys. Res.: Earth Surface}, 118(3), 1342-1355.

\bibitem[Silva et al (2006)]{Silva2006}
Silva, T. A. M., Bigg, G. R., and Nicholls, K. W. (2006), Contribution of giant icebergs to the Southern Ocean freshwater flux. {J. of Geophys. Res.: Oceans}, 111(C3).

\bibitem[Smith et al (2007)]{Smith2007}
Smith, K., B. Robison, J. Helly, R. Kaufmann, H. Ruhl, H., T. Shaw, and M. Vernet (2007), Free-drifting icebergs: Hotspots of chemical and biological enrichment in the Weddell Sea, {\it Science}, 317, 478–482.

\bibitem[Shepherd and Wingham (2007)]{Shepherd2007}
Shepherd, A., and D. Wingham (2007). Recent sea-level contributions of the Antarctic and Greenland ice sheets. {\it Science}, 315(5818), 1529-1532.

\bibitem[Stern et al (2014)]{Stern2014}
Stern, A. A., D. M. Holland,  P. R. Holland, A. Jenkins and J. Sommeria (2014), The effect of geometry on ice shelf ocean cavity ventilation: a laboratory experiment. {\it Experiments in Fluids}, 55(5), 1-19.

\bibitem[Stern et al (2015)]{Stern2015}
Stern, A.A., Johnson, E., Holland, D.M., Wagner, T.J., Wadhams, P., Bates, R., Abrahamsen, E.P., Nicholls, K.W., Crawford, A., Gagnon, J. and Tremblay, J.E. (2015), Wind-driven upwelling around grounded tabular icebergs. {\it J. of Geophys. Res.: Oceans}, 120(8), 5820-5835.

\bibitem[Stern et al (2016)]{Stern2016}
Stern, A. A., A. Adcroft, and O. Sergienko (2016), The effects of Antarctic iceberg calving?size distribution in a global climate model. {\it J. of Geophys. Res.: Oceans}, 121(8), 5773-5788.

\bibitem[Swope et al (1982)]{Swope1982}
Swope, W. C., H. C. Andersen, P. H. Berens, and K. R. Wilson (1982), A computer simulation method for the calculation of equilibrium constants for the formation of physical clusters of molecules: Application to small water clusters. {\it The Journal of Chemical Physics} 76, no. 1, 637-649.

\bibitem[Tournadre et al (2016)]{Tournadre2016}
Tournadre, J., N. Bouhier, F. Girard-Ardhuin, and F. Rémy (2016), Antarctic icebergs distributions 1992-2014. {\it J. Geophys Res: Oceans}.

\bibitem[Turnbull et al (2015)]{Turnbull2015}
Turnbull I.D., N. Fournier, M. Stolwijk, T. Fosnaes, D. McGonigal (2015), Operational iceberg drift forecasting in Northwest Greenland, {\it Cold Reg. Sci. Technol.} 110, 1-18

\bibitem[Unger (2014)]{Unger2014}
Unger, J. D., 2014. Regulating the Arctic Gold Rush:  Recommended Regulatory Reforms to Protect Alaska's ArcticEnvironment from Offshore Oil Drilling Pollution . {\it Alaska L. Rev}, 31

\bibitem[Vernet et al (2012)]{Vernet2012}
Vernet, M., et al. (2012), Islands of ice: Influence of free-drifting Antarctic icebergs on pelagic marine ecosystems, {Oceanography}, 25(3), 38–39

\bibitem[Weeks and Campbell  (1973)]{Weeks1973}
Weeks, W. F., and  W. J. Campbell (1973). Icebergs as a fresh-water source: an appraisal. {\it J. of Glaciol.}, 12(65), 207-233.

\bibitem[White et al  (2009)]{White2009}
White, L., A. Adcroft, and R. Hallberg (2009), High-order regridding–remapping schemes for continuous isopycnal and generalized coordinates in ocean models. {\it J. of Comp. Phys.}, 228(23), 8665-8692.

\end{thebibliography}

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/Regular_towards_coast}
\caption{ {Results of an ice-only LIISM simulation with no bonds between ice elements. Ice elements are initialized throughout the domain, as shown in panel (a). The elements are forced by an imposed westward ocean current of u=0.1m/s (no ocean model is used). Forces due to sea surface slope, atmospheric drag, Coriolis and sea ice drag are set to zero. The figure shows snapshots of ice element positions at time (a) t=0, (b) 200, (c) 400, (d) 600 days. The size of the dots shows the surface area (and interaction diameter) of each ice element. Land points are shown by black circles.}}
\end{center}
\label{fig:Regular_Uncoupled_sim}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/Tabular_and_regular_towards_coast_manual}
\caption{ {Results of an ice-only LIISM simulation using bonds between elements. Ice elements are initialized throughout the domain, as shown in panel (a). Three tabular icebergs are included, with 25, 16 and 4 elements respectively. The elements are forced by an imposed westward ocean current of u=0.1m/s (no ocean model is used). Forces due to sea surface slope, atmospheric drag, Coriolis and sea ice drag are set to zero. The figure shows snapshots of ice element positions at time (a) t=0, (b) 300, (c) 600, (d) 900 days. The size of the dots shows the surface area (and interaction diameter) of each ice element.  Bonds between ice elements are plotted in magenta. Land points are shown by black circles.}}
\end{center}
\label{fig:Uncoupled_sim}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/hex_intersecting_rectangles}
\caption{ {(a) Hexagonal elements are initialized in a staggered lattice as shown. Adjacent elements are bonded together. The centers of bonded elements are plotted in pink. The element bonds form equilateral triangles which give the larger structure rigidity. The black circles shows to the interactive length scales used in element interactions. (b) Intersection of a hexagonal element and an ocean grid. The colors indicate the fraction of the hexagon that lies in each grid cell. These fractions are used as weights to spread LIISM properties to the ocean grid (see text for more details).}}
\end{center}
\label{fig:Hex_intersections}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/ALE_z_static_shelf_solo_D_spread_mass_mass.png}
\caption{ {(a) Ocean bottom topography in MISOMIP configuration. (b) ice-shelf draft used in static shelf experiment. The ice draft is calculated from the ice mass in an ocean grid cell, which is found by spreading ice mass across ocean cells accounting for the size of each element (as explained in Section 2.3). (c) Same as in panel (b), except that the interpolation does not account for iceberg size, and instead treats elements as point masses.}}
\end{center}
\label{fig:ISOMIP_mass_and_topog}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Collapse_spread_area.png}
\caption{ {Snapshots of the fraction of ice cover in the LIISM tabular iceberg calving experiment. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. The dashed line in panel (c) shows the location of the vertical transects shown in Figures \ref{fig:Temperature_section_Collapse}} and \ref{fig:Velocity_section_Collapse}.}
\end{center}
\label{fig:Spread_area}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Collapse_temp_layers_x10.png}
\caption{ {Snapshots of vertical sections of ocean temperature at $x=58 $km in the LIISM tabular iceberg calving experiment. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. The position of the vertical transects is shown by the dashed lines in Figure \ref{fig:Spread_area}c.}}
\end{center}
\label{fig:Temperature_section_Collapse}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Collapse_temp.png}
\caption{ {Snapshots of the sea surface temperature in the LIISM tabular iceberg calving experiment. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. Regions with ice area fraction = 1 area plotted in white.}}
\end{center}
\label{fig:SST_Collapse}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Collapse_v_layers_x10.png}
\caption{ {Snapshots of vertical sections of meridional velocity at $x=58 $km in the LIISM tabular iceberg calving experiment. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. The position of the transects is shown by the dashed line in Figure \ref{fig:Spread_area}c.}}
\end{center}
\label{fig:Velocity_section_Collapse}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/ALE_z_static_shelf_solo_melt_m_per_year_sst_ustar_iceberg_calved.png}
\caption{ {Results of the tabular iceberg calving simulation 30 days after the iceberg calves. The three panels show snapshots of the (a) melt rate, (b) top-of-ocean temperature and (c) $u^{*}$ at the base of the ice shelf. Fields are only shown in regions where the ice area fraction is $\geq 0.8$.}}
\end{center}
\label{fig:Collapse_solo_melt}
\end{figure}
 \clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/drift_bond_temp_layers.png}
\caption{ {Results from the tabular iceberg calving experiment with and without iceberg bonds. The top row shows the fractional ice cover for the simulations (a) with and (b) without numerical bonds. The bottom row shows the corresponding vertical temperature section at $x=58 $km for the simulation (c) with and (d) without bonds. The location of the vertical transects in panels (c) and (d) are shown by the dashed lines in panels (a) and (b), respectively. All snapshots are taken at time t = 30 days. The simulations uses wind stress $\vec{\tau}=< 0.0, 0.05 >.$}}
\end{center}
\label{fig:Bond_vs_No_bond}
\end{figure}

 \clearpage
 \end{article}

 \end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%